### 2022-02-22 manuscript.Rmd
initial commit

---
title: "comparative_RNA-Seq_manuscript"
author: "Thomas_Nicholson"
date: "19/02/2022"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggtree)
library(treeio)
library(ape)
library(UpSetR)
library(svglite)
library(ROSE)
library(ggpubr)
source('~/bin/R/functions.R')

data_path = '~/bin/R/r_files/'
figure_path = '~/Google Drive/My Drive/Comparative_RNA-Seq/'
```

## UpsetR figure

```{r upsetr_tree}
#ML phylogenetic tree built using 16s rRNA for the Gammaproteobacteria
#The full tree was also built for all bacterial genera, but is not displayed
tree <- read.tree(paste0(data_path, 'upsetr.tree'))

  tbl_tree <- as_tibble(tree)

#list of extra Gammaproteobacteria that were 
#not included in the RNA-sed analysis
to_drop <- c("Azotobacter", "Marinobacter", "Pseudoalteromonas", "Agarivorans", 
             "Vibrio", "Alishewanella", "Aggregatibacter", "Mannheimia", 
             "Actinobacillus", "Xenorhabdus", "Providencia", "Proteus",
             "Pantoea", "Brenneria", "Lonsdalea", "Buchnera.", "Wigglesworthia", 
             "Sodalis", "Dickeya", "Citrobacter", "Plautiasymbiont", 
             "Shewanella", "Moritella", "Moraxella", "Psychrobacter", 
             "Methylomonas", "Cycloclasticus", "Methylococcus", "Francisella",
             "Pseudoxanthomonas", "Candidatus", "Plautia", "Methylophaga", 
             "Pasteurella", "Salinivibrio")

  sub_tree <- drop.tip(tree, to_drop)

    tbl_sub_tree <- as_tibble(sub_tree)

p <- ggtree(sub_tree) + 
  geom_tiplab(align = T) +
  xlim(0, 0.35)

p
if(FALSE){
ggsave(filename = paste0(figure_path, "SVG/upsetr_subset_tree.svg"), 
       plot = p, width = 8, height = 16)
}
p <- ggtree(tree) + 
  geom_tiplab(align = T) +
  xlim(0, 0.5)

p
if(FALSE){
ggsave(filename = paste0(figure_path, "SVG/upsetr_tree.svg"), 
       plot = p, width = 8, height = 16)
}
```

```{r upsetr_plot}
#matrix of bool values indicating for each RNA if it was found in a genus
load(paste0(data_path, "upsetSubsetPC.Rda"))
load(paste0(data_path, "upsetSubsetPredicted.Rda"))

#list of extra Gammaproteobacteria that were 
#not included in the RNA-sed analysis
genera_arranged <- c("Lysobacter", "Stenotrophomonas", "Xylella", "Xanthomonas", 
                     "Methylomicrobium", "Pseudomonas", "Acinetobacter", 
                     "Alteromonas", "Photorhabdus", "Yersinia", "Erwinia", 
                     "Edwardsiella", "Serratia", "Klebsiella", "Enterobacter", 
                     "Salmonella", "Shigella", "Escherichia")

upset_pred <- select_columns_by_list(upsetSubsetPredicted, genera_arranged)
upset_pc <- select_columns_by_list(upsetSubsetPC, genera_arranged)

UpSetR::upset(upset_pc, sets = colnames(upset_pc), 
              mb.ratio = c(0.55, 0.45), order.by = "freq", keep.order = T)
UpSetR::upset(upset_pred, sets = colnames(upset_pred), 
              mb.ratio = c(0.55, 0.45), order.by = "freq", keep.order = T)

if(FALSE){
svglite(filename=paste0(figure_path, "SVG/upsetr_pc.svg"))
UpSetR::upset(upset_pc, sets = colnames(upset_pc), 
              mb.ratio = c(0.55, 0.45), order.by = "freq", keep.order = T)
dev.off()
svglite(filename=paste0(figure_path, "SVG/upsetr_pred.svg"))
UpSetR::upset(upset_pred, sets = colnames(upset_pred), 
              mb.ratio = c(0.55, 0.45), order.by = "freq", keep.order = T)
dev.off()
}

```

```{r expressed_regions_frequency_plot}
#matrix of bool values indicating for each RNA if it was found in a genus
load(paste0(data_path, "upsetSubsetPC.Rda"))
load(paste0(data_path, "upsetSubsetPredicted.Rda"))

#list of extra Gammaproteobacteria that were not included in the RNA-sed analyis
genera_arranged <- c("Lysobacter", "Stenotrophomonas", "Xylella", "Xanthomonas", 
                     "Methylomicrobium", "Pseudomonas", "Acinetobacter", 
                     "Alteromonas", "Photorhabdus", "Yersinia", "Erwinia", 
                     "Edwardsiella", "Serratia", "Klebsiella", "Enterobacter", 
                     "Salmonella", "Shigella", "Escherichia")

upset_pred <- select_columns_by_list(upsetSubsetPredicted, genera_arranged)
upset_pc <- select_columns_by_list(upsetSubsetPC, genera_arranged)

pc_freq <- colSums(upset_pc)
pred_freq <- colSums(upset_pred)

pc_freq <- vector_to_dataframe(pc_freq, 'genera')
pred_freq <- vector_to_dataframe(pred_freq, 'genera')

pc_freq <- pc_freq %>% dplyr::rename(freq = vec) %>% mutate(group ='pc')
pred_freq <- pred_freq %>% dplyr::rename(freq = vec) %>% mutate(group ='pred')

freq_data <-pc_freq %>% bind_rows(pred_freq)

freq_data$genera <- factor(freq_data$genera, levels = unique(freq_data$genera))

p <- ggplot() +
  geom_bar(data = freq_data, 
           aes(y = genera, x = freq, group = group, fill = group), 
           stat = 'identity', position = 'dodge')

p

if(FALSE){
  ggsave(filename = paste0(figure_path, 'SVG/rnas_frequency.svg'), 
  plot = p, width = 8, height = 16)
}

```

## Random forest interpretation figure

```{r random_forest_frequency_plots}
load("~/bin/PhD/Chapter_4/chapter_4_files/predDat.Rda")
load("~/bin/PhD/Chapter_4/chapter_4_files/validatation2.Rda")

#select the desired columns from the predicted data and validatation data
probDat <- predDat %>% 
  select(probability, ID, group, srna.counts.2) %>% 
  bind_rows(validation2 %>% select(probability, ID, group, srna.counts.2))

#not sure if setting factors will break anything so using another data frame
plotDat <- probDat
plotDat$group <- factor(plotDat$group, 
                        levels = c('Positive Control', 
                                   'Predicted', 
                                   'Negative Control'))

#plot histogram of the probabilities
p <- ggplot() +
  geom_histogram(data = plotDat, 
                 aes(x = probability, 
                     y = ..density.., 
                     group = group, 
                     fill = group), 
                 binwidth = 0.02) +
  geom_vline(xintercept = 0.17) +
  geom_vline(xintercept = 0.5) 
p

if(FALSE){
  ggsave(filename = paste0(figure_path, "SVG/histogram_probabilities.svg"), 
         plot = p, width = 178, height = 155, units = "mm")
}

#get the cumulative counts of the number of alignments as probability increases
countsCumul <- cumulativeCounts(dists = probDat, 
                                smooth = F, 
                                target_column = 'probability')

##produces plot to use for figure showing probability results
p <- ggplot() +
  geom_line(data = countsCumul, aes(x= probability, 
                                    y = cumulative_prop,
                                    group = group, 
                                    colour = group),
            size = 1.5, 
            show.legend = F) +
  scale_y_continuous(trans = 'log10')
p + theme_classic()

if(FALSE){
  ggsave(filename = paste0(figure_path, "SVG/cumulative_probabilities.svg"), 
                           plot = p, width = 178, height = 155, units = "mm")
}
```

```{r random_forest_statistics}
load("~/bin/PhD/Chapter_4/chapter_4_files/predDat.Rda")
load("~/bin/PhD/Chapter_4/chapter_4_files/validatation2.Rda")

#select the desired columns from the predicted data and validatation data
probDat <- predDat %>% 
  select(probability, ID, group, srna.counts.2) %>% 
  bind_rows(validation2 %>% 
              select(probability, ID, group, srna.counts.2))

#not sure if setting factors will break anything so using another data frame
plotDat <- probDat
plotDat$group <- factor(plotDat$group, 
                        levels = c('Positive Control', 
                                   'Predicted', 
                                   'Negative Control'))
dat <- allScores(plotDat, 0.01, 'probability')

#get the FNR and PPV values at each threshold
if(FALSE){
  for(i in seq(0,1, by=0.01)){
    scores <- scoreProbabities(plotDat, 
                               threshold = i, 
                               target_column = 'probability')
    print(paste0(i, ': ', scores$fnr, ', ', scores$ppv))
  }
}

scores <- scoreProbabities(plotDat, 
                           threshold = 0.17, 
                           target_column = 'probability')
printListSubset(scores, 
                vec = c('ppv', 'fnr', 'pred_pos', 'pred_pct'), 
                startText = 'p > 0.17', round_val = 3)

scores <- scoreProbabities(plotDat, 
                           threshold = 0.5, 
                           target_column = 'probability')
printListSubset(scores, 
                vec = c('ppv', 'fnr', 'pred_pos', 'pred_pct', 'pc_pct'), 
                startText = 'p > 0.5', round_val = 3)

scores <- scoreProbabities(plotDat, 
                           threshold = 0.81, 
                           target_column = 'probability')
printListSubset(scores, 
                vec = c('ppv', 'fnr', 'pred_pos', 'pred_pct', 'pc_pct'), 
                startText = 'p > 0.81', round_val = 3)

```

```{r statistics_plots}
load("~/bin/PhD/Chapter_4/chapter_4_files/predDat.Rda")
load("~/bin/PhD/Chapter_4/chapter_4_files/validatation2.Rda")

#select the desired columns from the predicted data and validatation data
probDat <- predDat %>% 
  select(probability, ID, group, srna.counts.2) %>% 
  bind_rows(validation2 %>% 
              select(probability, ID, group, srna.counts.2))

#not sure if setting factors will break anything so using another data frame
plotDat <- probDat
plotDat$group <- factor(plotDat$group, 
                        levels = c('Positive Control', 
                                   'Predicted', 
                                   'Negative Control'))
dat <- allScores(plotDat, 0.01, 'probability')

ggplot(data = dat) + 
  geom_line(aes(x = threshold, y = ppv), color = 'blue') + 
  geom_line(aes(x = threshold, y = fnr), color = 'red')

roc.curve(response = plotDat$group[plotDat$group != 'Predicted'], 
          predicted = plotDat$probability[plotDat$group != 'Predicted'])

ggplot(data = dat) + 
  geom_line(aes(x = fnr, y = specificity), color = 'blue') + 
  geom_line(aes(x = fpr, y = sensitivity), color = 'red')

```

```{r repeat_of_cumulative_distributions}
#is there a difference between low scoring (p < 0.17 and high scoring p > 0.5) predicted rufs?

load("~/bin/Phd/Chapter_4/chapter_4_files/predDat.Rda")
load("~/bin/Phd/Chapter_4/chapter_4_files/validatation2.Rda")
##function written using max_dist as column name so each variable needs to  be renamed to this before using cumulativeDistribution()

featuresSelected <- predDat %>% 
  mutate(group = ifelse(probability > 0.75, 'high', ifelse(probability <= 0.17, 'low', 'drop'))) %>% 
  filter(group != 'drop') %>% select(group, distance, read.max.score, 
                                     cov.min.eval, z_max, motif.max.score,
                                     alifold_cov_score, mfe.score, 
                                     gc.score) %>% 
  bind_rows(validation2 %>% filter(group == 'Positive Control') %>% 
              select(group, distance, read.max.score, 
                                     cov.min.eval, z_max, motif.max.score,
                                     alifold_cov_score, mfe.score, 
                                     gc.score))

dat <- featuresSelected %>% dplyr::rename(max_dist = distance) %>% filter(max_dist <= 0.15)
distance.p <- cumulativeDistribution(dat, alternative = "two.sided", show.legend = F)
distance.p <- distance.p +
   labs(y = "Cumulative Proportion", x = "Evolutionary distance")

##known sRNAs and predicted RUFs are only selected if there is read depths. For a fair comparison, RINCs with read depths of 0 are removed.
dat <- featuresSelected %>% select(group, read.max.score) %>% dplyr::rename(max_dist = read.max.score) %>% filter(max_dist > 0)
reads.p <- cumulativeDistribution(dat, alternative = 'two.sided', show.legend = F)
reads.p <- reads.p +
   labs(y = "Cumulative Proportion", x = "Total reads")+
  scale_x_continuous(trans = "log10")
#reads.p

dat <- featuresSelected %>% mutate(max_dist = -log(cov.min.eval))
rscape.p <- cumulativeDistribution(dat)
rscape.p <- rscape.p +
   labs(y = "Cumulative Proportion", x = "Rscape covariance score")
# rscape.p

##none of the z scores are greater than 3, so the NA value of 10 is changed to 3 (then the negaive is taken for the plot)
dat <- featuresSelected %>% mutate(max_dist = ifelse(z_max == 10, -3, -z_max)) %>% select(group, max_dist)
z.p <- cumulativeDistribution(dat, show.legend = F)
z.p <- z.p +
   labs(y = "Cumulative Proportion", x = "Alifold z-score (negative energy)")

# z.p
##selected a window where the values are easier to visualise. This has removed 16 known sRNAs and 9 predicted RUFs
dat <- featuresSelected %>% dplyr::rename(max_dist = motif.max.score) %>% filter(max_dist < 1000)
motif.p <- cumulativeDistribution(dat)
motif.p <- motif.p +
   labs(y = "Cumulative Proportion", x = "Motif score")

dat <- featuresSelected %>% mutate(max_dist = -alifold_cov_score) 
alifold.cov.p <- cumulativeDistribution(dat)
alifold.cov.p <- alifold.cov.p +
   labs(y = "Cumulative Proportion", x = "Alifold covariance score")

dat <- featuresSelected %>% mutate(max_dist = -mfe.score)
mfe.p <- cumulativeDistribution(dat, alternative = 'two.sided')
mfe.p <- mfe.p +
   labs(y = "Cumulative Proportion", x = "MFE score (negative energy)")

dat <- featuresSelected %>% arrange(gc.score) %>% mutate(gc.score = round(gc.score))

highCounts <- dat %>% filter(group == "high") %>% group_by(gc.score) %>% summarise(count = n()) %>% arrange(gc.score) %>% ungroup() %>% tidyr::complete(gc.score = seq(from = 0, to = 100, by = 1), fill = list(count = 0))
lowCounts <- dat %>% filter(group == "low") %>% group_by(gc.score) %>% summarise(count = n()) %>% arrange(gc.score) %>% ungroup() %>% tidyr::complete(gc.score = seq(from = 0, to = 100, by = 1), fill = list(count = 0))

highTotal <- dat %>% filter(group == "high") %>% nrow()
lowTotal <- dat %>% filter(group == "low") %>% nrow()

highGC <- zoo::zoo(highCounts$count)
lowGC <- zoo::zoo(lowCounts$count)

smoothHigh <- zoo::rollapply(highGC, width = 10, by = 1, FUN = mean, align = "center", partial = T) 
smoothLow <- zoo::rollapply(lowGC, width = 10, by = 1, FUN = mean, align = "center", partial = T)

smoothHigh <- as.data.frame(smoothHigh) %>% mutate(x = row_number() -1) %>% mutate(group = "high") %>% dplyr::rename(y = smoothHigh) %>% mutate(y = y/highTotal)
smoothLow <- as.data.frame(smoothLow) %>% mutate(x = row_number() -1) %>% mutate(group = "low") %>% dplyr::rename(y = smoothLow) %>% mutate(y = y/lowTotal)

pcCounts <- dat %>% filter(group == "Positive Control") %>% group_by(gc.score) %>% summarise(count = n()) %>% arrange(gc.score) %>% ungroup() %>% tidyr::complete(gc.score = seq(from = 0, to = 100, by = 1), fill = list(count = 0))
pcTotal <- dat %>% filter(group == "Positive Control") %>% nrow()
pcGC <- zoo::zoo(pcCounts$count)
smoothPC <- zoo::rollapply(pcGC, width = 10, by = 1, FUN = mean, align = "center", partial = T) 
smoothPC <- as.data.frame(smoothPC) %>% mutate(x = row_number() -1) %>% mutate(group = "Positive Control") %>% dplyr::rename(y = smoothPC) %>% mutate(y = y/pcTotal)

smoothGC <- smoothHigh %>%  bind_rows(smoothLow, smoothPC)

gc.p <- ggplot() +
  geom_path(data = smoothGC, aes(x = x, y = y, group = group, color = group), size = 1, show.legend = FALSE)  + labs(y = "Proportion", x = "GC percentage")  + theme_classic()

all.p <- ggarrange(distance.p, reads.p, mfe.p, z.p, rscape.p, alifold.cov.p,  motif.p, gc.p + rremove("x.text"),
          labels = LETTERS[1:9],
          ncol = 3, nrow = 3)

all.p

if(FALSE){
ggsave(filename = paste0(figure_path, "SVG/separated_distributions.svg"), plot = all.p, width = 450, height = 307, units = "mm")
}

```